---
title: "riskyData-Welcome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskyData-Welcome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to data importing

### Welcome to riskyData!

The package `riskyData` has been developed for the Evidence and Risk colleagues to cleanly and reliably import data from the Environment Agencies API.

## Loading the package

The `riskyData` package is part of a group of tools called `flode`. To load all these packages, the easiest way is to use;

```{r flode, eval = FALSE}
library(flode)
```

However, if you only require just the `riskyData` package you can just use;

```{r setup}
library(riskyData)
```

## Example I - List all sites in API

The `loadAPI()` function forms the basis of `riskyData.` Most operations relate in some manner to this. Running the code with no parameters or `ID = NULL` or `ID = "all"` will return the station list

This is not just limited to WISKI sites, some water quality data are also present. This functionality will be expanded upon in future versions.

Note this can take some time to load

```{r loadAll}
loadAPI()
```

If you wish to constrain these data to NRFA or WISKI sites only add "nrfa" or "wiski" to the `ID` argument. There's also an additional API that this function can connect to that contains real time data, for real time data use "flood" and for tidal sites use "tidal".

Using these in the `obsProperty` argument you can filter your results to certain types of gauges. With this you are limited to;

```{r observedPs, echo = FALSE}
Parameter <- c('Flow', 'Stage/Level', 'Rainfall', 'Groundwater Level', 'Ammonium',
          'Disolved Oxygen', 'Conductivity', 'pH', ' Temperature', 'Turbidity',
          'Nitrate', 'Chlorophyll', 'Salinity', 'Blue Green Algae',
          'Fluorescent Dissolved Organic Matter')

observedProperty <- c('waterFlow', 'waterLevel', 'rainfall', 'groundwaterLevel', 'ammonium', 'dissolved-oxygen', 'conductivity', 'ph', 'temperature', 'turbidity', 'nitrate', 'chlorophyll', 'salinity', 'bga', 'fdom')

Unit <- c('m3S-1', 'm', 'mm', 'mAOD', 'mg/L', 'mg/L or %', '$\\mu /S$ cm', '', 'C',
          'NTU', 'mg/L (as N)', '$\\mu /L$', 'g/L', 'RFU', 'RFU')
df <- data.frame(Parameter, observedProperty, Unit)

knitr::kable(df, caption = "Available datapoint options")
```

To load all data you can also use;

```{r alternateLoads, eval = FALSE}
# Note code not run
loadAPI(ID = "all") # does the same as loadAPI()
```

If we wished to load all the flow sites with WISKI data available we can use;

```{r WISKI}
loadAPI(ID = "wiski", obsProperty = "waterFlow")
```

Or if we wished to load the sites that are used within the National River Flow Archive (NRFA);

```{r NRFA}
loadAPI(ID = "nrfa")
```
To view sites that have realtime data available use;

```{r flood}
loadAPI(ID = "flood")
```

Finally for tidal sites use;

```{r tidal}
loadAPI(ID = "tidal")
```
## Example II - Find nearby gauges

Assuming we had a site of interest and wished to see the available nearby gauges. We can search for sites with coordinates within a certain range.

Using the following arguments we can search the api for nearby data; `easting`, `northing`, `lat`, `long`, and `dist.`

When we have easting and northing coordinates of 378235 and 276165 respectively and wish to find NRFA gauges within 30km;

```{r distance1}
loadAPI(easting = 378235 , northing = 276165, dist = 30, ID = "nrfa")
```

Alternatively, find all groundwater data within 10km of lat-long coordinates; 52.58302, -0.40656

```{r distance2}
loadAPI(lat = 52.58302, long = -0.40656, dist = 100, obsProperty = "groundwaterLevel")
```

**Note** the function will not allow you to mix easting-northing with lat-long.

```{r distance3}
loadAPI(easting=247613, lat = 50.5, long = -4.15, dist = 10, ID = "nrfa")
```

Once you have identified a dataset of interest, such as Bewdley, we can inform the API to return all the available datasets assigned to the WISKI ID;

## Example III - Known WISKI ID

Say we wish to see what data are available for a given site, in this case Bewdley (WISKI 2001)

We can now use the WISKI ID in the `ID` argument in `loadAPI()`

```{r datasets}
loadAPI(ID = '2001') #! It's good practice to input the ID as a character string
```

As you can see there are 3 level and 4 flow series available. Of these there are are 2 time steps or periods available. Daily data are 86400 seconds and 15 minute are 900 seconds. The 15 minute data are of type "instantaneous" - these data are used in our flood models. For both flow and stage there are daily minimum and maximum values, for flow there is also a daily mean available.

We wish to now drill into the instantaneous data a bit more.

## Example IV - using the datapoints argument

The datapoints argument gives us a lot of control as to what we wish to pull from the API.

```{r table, echo = FALSE}
datapoints <- c('earliest', 'latest', 'day', 'range', 'standard', 'all')
info <- c('Returns the first available timestep',
          'Returns the last available timestep',
          'Provides a full days worth of data, uses the from argument to specify the date',
          'Exports all data between to dates, uses from and to arguments',
          'Default extraction of up to 200,000 rows',
          'Increases the data pull to cover everything available')
df <- data.frame(datapoints, info)

knitr::kable(df, caption = "Available datapoint options")

```

To find the first time step of the instantaneous series we need to include more arguments; `measure`, `period`, `type`, and `datapoints` .

All data are exported in the `HydroImport` class, with this gauge metadata is pulled simultaneously with the observed data. More details on these classes can be seen in VIGNETTE XXXX

```{r earliest, eval = FALSE}
loadAPI(ID = '2001', 
        measure = 'flow',
        period = 900,
        type = 'instantaneous',
        datapoints = 'earliest')
```

```{r earliestBodge, echo = FALSE}
data(bewdley)

a <- bewdley$clone()
a$data <- a$data[1,]
a
```

For the latest we can change the text of the `datapoints` argument to **'latest'**

```{r latest, eval = FALSE}
loadAPI(ID = '2001', 
        measure = 'flow',
        period = 900,
        type = 'instantaneous',
        datapoints = 'latest')
```

```{r latestBodge, echo = FALSE}
data(bewdley)

b <- bewdley$clone()
b$data <- tail(b$data, 1)
b
```

Now say we wished to look at a specific day **'2022-06-11'**. To do this we will change the `datapoints` argument and add the `from` one. We will also plot the data for this;

```{r day, eval = FALSE}
dt <- loadAPI(ID = '2001', 
              measure = 'flow',
              period = 900,
              type = 'instantaneous',
              datapoints = 'day',
              from = '2022-06-11')

dt$hydroYearDay()$plot(wrap = FALSE)
```

```{r dayBodge, echo = FALSE, fig.width = 7, fig.height = 5}
data(bewdley)

c <- bewdley$clone()
c$data <- c$data[dateTime >= '2022-06-11 00:00' & dateTime <= '2022-06-12 00:00',]

c$hydroYearDay()$plot(wrap = FALSE)
```

If we wished to extend this we can use the **'range'** option in `datapoints`. For this we a beginning and end date, these are expressed with `from` and `to` arguments

```{r range, eval = FALSE}
dt <- loadAPI(ID = '2001', 
              measure = 'flow',
              period = 900,
              type = 'instantaneous',
              datapoints = 'range',
              from = '2019-10-01 00:00',
              to = '2020-03-30 00:00')

dt$hydroYearDay()$plot(wrap = FALSE)
```

```{r rangeBodge, echo = FALSE, fig.width = 7, fig.height = 5}
data(bewdley)

d <- bewdley$clone()
d$data <- d$data[dateTime >= '2019-10-01 00:00' & dateTime <= '2020-03-30 00:00',]

d$hydroYearDay()$plot(wrap = FALSE)
```

To pull all the data we can use the "all" option.

**Note** this can take a long time as there is a lot of data!

```{r all, eval = FALSE}
loadAPI(ID = '2001', 
        measure = 'level',
        period = 900,
        type = 'instantaneous',
        datapoints = 'all')
```

```{r allBodge, echo = FALSE}
data(bewdley)
bewdley
```
