---
title: "riskyData-Ratings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskyData-Ratings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Ratings in riskyData

Stage discharge rating tables can be added and be applied to data in the 
`HydroImport` containers.

Let's download some stage data;

```{r load, echo = FALSE, messages = FALSE}
library(riskyData)
```

```{r stage, message = FALSE, warning = FALSE, results = 'hide'}
bewdleyStage <- loadAPI(ID = "2001",
                        measure = "level",
                        period = 900,
                        type = "instantaneous",
                        datapoints = "range",
                        from = '2020-10-01 00:00', 
                        rtExt = TRUE)


```

To access the rating use; 

```{r stage1}
bewdleyStage$rating
```

As you can see the imported data doesn't yet have any rating. To add one we can 
use the `addRating()` method. This method has four arguments that need to be 
filled (C, A, B, and max).

```{r addRating}
bewdleyStage$addRating(C = c(43.3123, 43.3123, 47.6218, 0.000668, 0.000668),
                       A = c(-0.171831, -0.171831, -0.117731, -8.604474, -8.604474),
                       B = c(1.44155, 1.44155, 1.38813, 5.139, 5.139),
                       max = c(0.359, 0.504, 5.23, 5.55, 999))
```

This will now be displayed whenever you call the `HydroImport` container or 
using `$rating`;

```{r callUp}
bewdleyStage
```

To apply the rating to the stage data use the `rateFlow()` method;

```{r applyRating}
bewdleyStage$rateFlow()
```

To set the outputs in a `HydroImport` container, set the full argument to `TRUE`

```{r applyRating2}
bewdleyStage$rateFlow(full = TRUE)
```

## Rated stage

We can also use the same ratings to convert flow to stage;

```{r flowStage}
## Load flow data
data("bewdley")

## Set a window to make data comparable
bewdley$window(start = "2020-10-01 00:00", end = "2022-10-01 00:00", export = "snip")
bewdleyStage$window(start = "2020-10-01 00:00", end = "2022-09-30 23:45:00", export = "snip")

## Add the rating to Bewdley
bewdley$addRating(C = c(43.3123, 43.3123, 47.6218, 0.000668, 0.000668),
                  A = c(-0.171831, -0.171831, -0.117731, -8.604474, -8.604474),
                  B = c(1.44155, 1.44155, 1.38813, 5.139, 5.139),
                  max = c(0.359, 0.504, 5.23, 5.55, 999)
                  )

## Calculate rated stage
ratedStage <- bewdley$rateStage()
ratedStage
```

We can compare against the previously imported stage data.

```{r, echo = FALSE, messages = FALSE, warning = FALSE}
# library(ggplot2)
# bewdleyStage$data$ratedStage <- ratedStage$stage
```

```{r, echo = FALSE}
# ggplot(bewdleyStage$data, aes(x = dateTime, y = value)) +
#   geom_line() + 
#   geom_line(aes(y = ratedStage), colour = "red") +
#   xlab("Date Time") +
#   ylab("Stage (m)")
```
