---
title: "riskyData-Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskyData-Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE, warnings = FALSE}
library(riskyData)
library(ggplot2)
```


# Introduction to data visualisation in riskyData

### Plotting with riskyData!

The package `riskyData` has very simple methods for plotting hydrometric data, 
simply use the `$plot()` function. The metadata that is associated with the 
imported data will quickly specify the types of plots required. This limits the 
amount of additional arguments required.

All plots are generated with ggplot2. This is an implementation of Leland 
Wilkinson's Grammar of Graphics — a general scheme for data visualization which 
breaks up graphs into semantic components such as scales and layers and are 
stored in a list. Their positions in the list determine the plotting order when 
generating the graphical output. ggplot2 serves as a replacement for the base 
graphics and contains a number of defaults for web and print display of common 
scales. A critical element of `riskyData` is the OOP that underpins the package,
using ggplot further adds to this functionality as plotting objects can be 
stored and modified, rather than direct exporting.

Plots are currently associated with the formats useful for E&R forecasting, 
additional themes will be added soon.

## Loading the data

For the purpose of this we will import the Bewdley flow and Chesterton rain 
gauge data sets

All plotting is organised around hydrological years. Use the `$hydroYearDay()`
function in any pipelines.

```{r data}
## Load data
data(bewdley)
data(chesterton)

## Convert to hydrological year and day
bewdley$hydroYearDay()
chesterton$hydroYearDay()
```


## Standard plots

For flow and level stations, when you apply the `$plot()` function the plots are
faceted by hydrological year and default to a line based geometry.

```{r baseplot_flow, fig.width = 7, fig.height = 12}
bewdley$plot()
```

For rainfall stations, when you apply the `$plot()` function the plots are
also faceted by hydrological year but default to a column based geometry.

```{r baseplot_rain, fig.width = 7, fig.height = 12}
chesterton$plot()
```

## Viewing the whole series

For all types of gauges, you can view an unfaceted plot if you apply the wrap 
argument as FALSE. The basic geometries of lines for level and flow plots and 
bars/columns for rainfall sites will be retained.

If you wish to remove the title, set the title argument to FALSE.

```{r baseplot_unwrap, fig.width = 7, fig.height = 6}
bewdley$plot(wrap = FALSE, 
             title = FALSE)
```


## Cumulative rainfall

It is sometimes useful to view rainfall using a cumulative sum. With `riskyData`
 these can be applied unwrapped or wrapped. If wrap is set as TRUE, the 
cumulative plots are calculated for each hydrological year.

To set a cumulative plot set cumul to TRUE.
 
```{r cumul_unwrap, fig.width = 7, fig.height = 6}
chesterton$plot(wrap = FALSE, 
                cumul = TRUE)
```

To calculate and plot cumulative rainfall by hydrological year set wrap to TRUE.

```{r cumul_wrap, fig.width = 7, fig.height = 12}
chesterton$plot(wrap = TRUE, 
                cumul = TRUE)
```

## Customising your plots

As the plots can be stored as objects, this enables us to add extra information 
to them. As mentioned above the objects are stored as lists, to add extra 
details we can modify the objects as you would any other ggplot2 object.

In the example below I want to add threshold lines to the bewdley flow data that
 is not faceted. Additionally I wish to set the theme to linedraw and facet the 
plot so that there are 4 columns.


```{r threshold, fig.width = 7, fig.height = 12}
## Set threshold examples
thresholds <- c(300, 400, 500)

## Store plot as object
plot <- bewdley$plot(wrap = FALSE) 

## Modify plot
plot + geom_hline(yintercept = thresholds, 
                  colour = 'red',
                  linetype = 'dashed',
                  linewidth = 0.8) + 
  facet_wrap(~hydroYear, 
             scales = 'free_x', 
             ncol = 3) +
  theme_linedraw()
```

If we wished to plot the peaks detected by the `$findPeaks()` function.

```{r peaks, fig.width = 7, fig.height = 12}
bewdley$findPeaks()

plot <- bewdley$plot(wrap = FALSE)
plot + geom_point(data = bewdley$peaks, 
                  inherit.aes = FALSE,
                  aes(x = dateTime, y = value), 
                  colour = 'red') +
  facet_wrap(~hydroYear, scales = 'free_x')
```
