---
title: "riskyData-Containers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskyData-Containers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Containers in riskyData

If you have seen the previous vignette `riskyData-Welcome` you might have recognised that when data are pulled from the API they aren't in a conventional format. When data are scraped using the `loadAPI()` function , gauge metadata are also pulled simultaneously.

Using the `bewdley` dataset we can investigate;

```{r setup}
library(riskyData)
data(bewdley)
bewdley
```

When the dataset is printed there are 2 sections: **private** and **public**. These are both grouped into a container called bewdley, and have a class of `HydroImport` and `R6`.

```{r r6}
class(bewdley)
```

Using R6 allows you to define private fields and methods, in addition to the public ones. What private means is that fields and methods can only be accessed within the class, and not from the outside. Whereas with public, you can interact and modify the fields.

R6 is an implemention of encapsulated object-oriented programming for R, and is a simpler, faster, lighter-weight alternative to R\'s built-in reference classes. This style of programming is also sometimes referred to as classical object-oriented programming.

Some features of R6:

-   R6 objects have reference semantics.

-   R6 cleanly supports inheritance across packages.

-   R6 classes have public and private members.

In contrast to R\'s reference classes, R6 is not built on the S4 class system, so it does not require the *methods* package. Unlike reference classes, R6 classes can be cleanly inherited across different packages.Public data

## Public members

Public members are accessible from outside the class and can be used and modified directly.

To call the raw data in `riskyData` you can call it with `$data` this does add one level of complexity against a normal data frame but the benefits outstrip the negatives.

To call up the data on the bewdley dataset use;

```{r public1}
bewdley$data
```

As with a normal dataframe we can interact with functions from outside of the `riskyData` package;

```{r public2}
mean(bewdley$data$value, na.rm = TRUE)
max(bewdley$data$value, na.rm = TRUE)
min(bewdley$data$value, na.rm = TRUE)
with(bewdley$data, plot(x = dateTime, y = value, type = 'l'))
```

## Private data

Private members are only accessible from within the class, and they are encapsulated to ensure data integrity.

All the gauge metadata are stored within the private section, with this you cannot directly interact with or edit these data. For example let's say I wished to change the catchment area of the bewdley dataset

```{r privateError, eval = FALSE}
bewdley$start <- now()
```

```{r privateError1, echo = FALSE}
cat("Error in bewdley$start <- now() : cannot add bindings to a locked environment\n")
```

The data are protected, this ensures that they can be used at all times in other dependent functions. There is only one way to interact with the private metadata and that is through functions built into the container.

## Active Bindings

Active bindings are special methods that allow you to compute values on-the-fly when accessing an attribute.

## Container functions

Functions specific to the data stored within R6 containers are called methods. To use these you interact with them in a different manner to how you normally do in R. They can all be applied using the `$` operator, meaning that you don't have to encapsulate an object within parenthesis.

Do find all the methods available to you use `$methods()` after the object name. For example;

```{r methods}
bewdley$methods()
```

Using methods inbuilt to the containers, you can interact with the private metadata;

```{r metadataInteract}
# Return NRFA details
bewdley$nrfa()

# Return gauge coordinate data
bewdley$coords()

# Return all the metadata
bewdley$meta()
```

More details on methods will be covered in the methods vignette.

## Inherited Members

Inherited members are methods and fields inherited from the parent class if your R6 class inherits from another.

When data are imported with the `loadAPI()` function, the container used is `HydroImport`. This is the parent class. If an aggregation method applied, it can fundamentally change the data structure. For this reason a child class was developed called `HydroAggs`. Most functionality is inherited from the `HydroImport` parent class, however some methods had to be amended to ensure that they would still work.

To generate a `HydroAggs` dataset we can use the `$dataAgg()` function. Here we will calculate the hourly maximum data from the bewdley dataset and export it as class `HydroAggs`;

```{r HydroAggs1, eval = FALSE}
bewdley$dataAgg(type = "hourly", method = "max")
```

```{r HydroAggs2, echo = FALSE, message = FALSE}
dt <- bewdley$dataAgg(type = "hourly", method = "max")
```

```{r HydroAggs3, echo = FALSE}
dt
```

Though there quite a lot of changes under the hood the output looks very similar to class `HydroImport`. The public data is different as we have applied an aggregation, with the private metadata the start and end times of the series differ. One of the key changes is the modifications line, all the containers used in `riskyData` track changes made to the data, in this case it indicates that an "hourly max" aggregation has been applied.
