---
title: "riskyData-Ratings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskyData-Ratings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Ratings in riskyData

Stage discharge rating tables can be added and be applied to data in the 
`HydroImport` containers.

Let's download some stage data;

```{r load, echo = FALSE, messages = FALSE}
library(riskyData)
```

```{r stage, eval = FALSE}
bewdleyStage <- loadAPI(ID = "2001",
                        measure = "level",
                        period = 900,
                        type = "instantaneous",
                        datapoints = "range",
                        from = '2022-10-01 00:00', 
                        rtExt = TRUE)
```

```{r stage2, echo = FALSE, messages = FALSE, warning = FALSE}
bewdleyStage <- loadAPI(ID = "2001",
                        measure = "level",
                        period = 900,
                        type = "instantaneous",
                        datapoints = "range",
                        from = '2022-10-01 00:00', 
                        rtExt = TRUE)

bewdleyStage
```

To access the rating use; 

```{r stage3}
bewdleyStage$rating
```

As you can see the imported data doesn't yet have any rating. To add one we can 
use the `addRating()` method. This method has four arguments that need to be 
filled (C, A, B, and max).

```{r addRating}
bewdleyStage$addRating(C = c(43.3123, 43.3123, 47.6218, 0.000668, 0.000668),
                       A = c(-0.171831, -0.171831, -0.117731, -8.604474, -8.604474),
                       B = c(1.44155, 1.44155, 1.38813, 5.139, 5.139),
                       # max = c(-0.171831, 0.359, 0.504, 5.23, 5.55)
                       max = c(0.359, 0.504, 5.23, 5.55, 999)
                       )
```

This will now be displayed whenever you call the `HydroImport` container or 
using `$rating`;

```{r callUp}
bewdleyStage
```

To apply the rating to the stage data use the `rateFlow()` method;

```{r applyRating}
bewdleyStage$rateFlow()
```

To set reatin the outputs in a `HydroImport` container, set the full argument 
to `TRUE`

```{r applyRating2}
bewdleyStage$rateFlow(full = TRUE)
```
